---
title: "Project 1"
author: "Mar"
date: "31 de julio de 2017"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, fig.path='Figs/')
```

# Project 1 of the Reproducible Research cours

## Loading and preprocessing the data

```{r}
setwd("C:/Users/Mar/Documents/formations/MOOCs/Coursera_DataScience/05_reproducible_research/w2_project/")

data<-read.csv('activity.csv', header=TRUE)

str(data)

data$date<-as.Date(as.character(data$date), format="%Y-%m-%d")
```


## Obtaining the mean total number of steps taken per day

```{r}
library(dplyr)
df<-tbl_df(data)
str(df)
per_date<-group_by(df, date)

# Obtain the total number of steps per day
summ<-summarise(per_date, a_sum=sum(steps, na.rm=TRUE))
#Make a histogram
hist(summ$a_sum, main="Mean number of steps per day")
#Calculate the mean and median of the total number of steps per day
mean_steps<-mean(summ$a_sum)
median_steps<-median(summ$a_sum)
```

The mean number of steps per day is `r mean_steps`.
The median nunmber of steps per day is `r median_steps`.


## Average daily activity pattern

```{r}
steps_int<-with(df, tapply(steps, interval, mean, na.rm=T))
plot(names(steps_int), steps_int, type="l", xlab="5-min Interval", ylab="Average steps")

max_int<-max(steps_int)
name_max_int<-names(steps_int[steps_int=max_int])
```

The 5-minute interval corresponding to the maximum number of steps is `r name_max_int`.


## Imputing missing values

```{r}
#Total number of missing data
miss_tot<-sum(is.na(df))

# New dataset with filled na
df_new<-df
miss<-df$interval[is.na(df_new$steps)]
index<-sapply(miss, function(x) which(names(steps_int)==x))
df_new$steps[is.na(df_new$steps)]<-steps_int[index]

# Obtain the new total number of steps per day
per_date_new<-group_by(df_new, date)
summ_new<-summarise(per_date_new, a_sum_new=sum(steps, na.rm=TRUE))
#Make a histogram
hist(summ_new$a_sum_new, main="Mean number of steps per day after filling NA")
#Calculate the mean and median of the total number of steps per day
mean_steps_new<-mean(summ_new$a_sum_new)
median_steps_new<-median(summ_new$a_sum_new)

```

The total number of missing values is `r miss_tot`.
The mean number of steps per day after filling the NA values is `r mean_steps_new`, while the original one was `r mean_steps`. Fillinig the missing data has increased the mean value per day. 
The median nunmber of steps per day after filling the NA values is `r median_steps_new`, while the orginale one was `r median_steps`. In this case, both values ar the same


## Differences in activity patterns between weekdays and weekends

```{r}
library(ggplot2)
# New variable 'weekday'
wd<-weekdays(df_new$date)
wkd<-wd
wkd[wd=="sábado" | wd=="domingo"]<-"weekend"
wkd[wd!="sábado" & wd!="domingo"]<-"weekday"

df_new$weekday<-as.factor(wkd)

steps_sum<-group_by(df_new, interval, weekday)
new_sum<-summarise(steps_sum, a_mean=mean(steps, na.rm=TRUE))

ggplot(new_sum, aes(interval, a_mean, col=weekday)) + facet_grid(weekday~.)+geom_line()+ylab("Mean number of steps")

```

---
output:
  md_document:
    variant: markdown_github
---
